\documentclass[11pt]{cdio}
%% Setup citation system
\usepackage[backend=biber, date=short, maxcitenames=2, bibencoding=utf8, style=apa]{biblatex}
\DeclareLanguageMapping{american}{american-apa}% needed for style=apa
%\renewcommand*{\bibfont}{\referencefont}
\setlength{\bibitemsep}{\baselineskip}
\setlength{\bibhang}{0pt}

\addbibresource{references.bib}% file with reference library

\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}
\usepackage{fontawesome7}
\usepackage{booktabs}

\graphicspath{{graphics/}{Graphics/}{include/}}%You can store your images in these locations

\title{Using Pull Requests to Make Collaboration Visible in CDIO Project-Based Courses}

\footer{Proceedings of the $22^\text{nd}$ International CDIO Conference, hosted by\\University of Liverpool, Liverpool, United Kingdom, June 22--25, 2026}

\setuphyperref{}%enable clickable links

\usepackage{enumitem}
\newlist{romanenumerate*}{enumerate*}{1}
\setlist[romanenumerate*]{
  label=(\roman*),
  itemjoin={{; }},
  itemjoin*={{; and }},
  before=\leavevmode,
  after={{.}}
}

\begin{document}
\makecdiotitle{
  \author{Helga Ingimundardóttir}
  \affiliation{Industrial Engineering, University of Iceland}
  % https://apastyle.apa.org/style-grammar-guidelines/capitalization/title-case
}

\section{Abstract}
Assessing collaboration, individual accountability, responsiveness to feedback, and reproducibility remains a persistent challenge in open-ended, team-based project courses. Final deliverables often mask how work evolved, how responsibilities were shared, and how decisions were negotiated. This paper presents an assessment design in which pull requests (PRs), a mechanism for reviewing and integrating changes to shared project work, serve as the central mechanism for making collaborative practices visible and assessable within CDIO-aligned project-based courses. Rather than treating GitHub as a programming topic or tooling exercise, the approach positions PRs as assessment infrastructure that binds technical change to process evidence through review discussions, explicit responses to comments, and documented handover.
The implementation spans two intentionally sequenced courses in Industrial Engineering. A second-year course introduces structured collaboration practices through short, skills-focused project cycles. A subsequent advanced course assumes GitHub fluency and organizes all project work within a single persistent repository maintained across the semester. Within this environment, PRs function as the primary unit of feedback, peer review, and evaluation. Students make their reasoning, revisions, and communication visible through routine workflow activity, and unresolved feedback carries forward across project cycles into capstone evaluation, enabling cumulative rather than fragmented assessment.
Experience across multiple offerings shows that PR-centered assessment improves the scope and specificity of peer review, strengthens documentation and reproducibility practices, and fosters shared responsibility for evolving team artifacts. Students initially expressed concern about perceived overhead, but review quality, documentation habits, and disciplined handover improved markedly as the workflow became routine. Graduating students later identified PR-based communication and accountability as directly transferable to professional settings.
The paper highlights key design decisions, trade-offs, and sequencing choices, offering practical and transferable guidance for educators seeking assessment methods that make collaboration, responsiveness, and reproducibility genuinely visible in CDIO project-based courses.

\section{Keywords}
Pull requests, GitHub Classroom, assessment infrastructure, team-based learning, professional competence, CDIO standards: 5, 7, 8, 11

% The implementation primarily addresses CDIO Standards 7, 8, and 11. Pull-request-based workflows integrate disciplinary, interpersonal, and professional skills within authentic project activity (Standard 7), while review-driven iteration and peer engagement support active, experiential learning (Standard 8). Assessment is grounded in observable artifacts that make individual contribution, collaboration quality, and responsiveness to feedback visible within team outputs (Standard 11). The approach additionally supports Design–Implement experiences (Standard 5) through sustained, iterative project work enabled by persistent repositories.

%%%%%%% DO NOT REMOVE %%%%%%%%%%%%%%%
\newpage{}% Title, authors, abstract, and keywords go on the first page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{include/body}

\section{ACKNOWLEDGEMENTS}
The author received no financial support for this work. The author thanks the students whose sustained engagement and willingness to adopt a demanding collaboration and assessment approach made this implementation possible.

%\section{REFERENCES}
\setlength\bibitemsep{2pt}% Word template is single spaced, but it looks like there is extra space between items in the official implementation --foley

\printbibliography{}

\section{Biographical Information}
\textbf{\textit{Helga Ingimundardóttir}} is an Assistant Professor of Industrial Engineering at the University of Iceland. Since 2023, her teaching and research have focused on optimization, mathematical modeling, data science, artificial intelligence, and industry-engaged learning in data-intensive contexts. She is a member of the Teaching Academy of the public universities in Iceland (2024).
Helga brings extensive industry experience across research and applied AI: as a Research Scientist at \emph{deCODE Genetics}, she designed long-read sequencing analysis pipelines and co-authored papers in \emph{Nature Genetics} and \emph{Genome Biology}; as a Data Scientist at \emph{CCP Games}, she helped develop real-time recommendation services for \emph{EVE Online}; as Head of AI Research at \emph{Travelshift}, she led large-scale trip-planning optimization; and earlier, as a Computational Engineer in R\&D at \emph{Valka} (now JBT Marel), she developed X-ray-based optimization methods for automated fish processing.
She holds a PhD and MSc in Computational Engineering and a BSc in Mathematics from the University of Iceland, as well as a graduate diploma in Teaching Studies for Higher Education.

\subsection{Corresponding author}
\CorrespondingAuthorBlock{
Helga Ingimundardóttir\\
University of Iceland\\
School of Engineering and Natural Sciences\\
Faculty of Industrial Engineering, Mechanical Engineering and Computer Science\\
VR-II, Hjarðarhagi 6\\
107 Reykjavík, ICELAND\\}
{helgaingim@hi.is}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: xetex
%%% End:
