\documentclass[11pt]{cdio}
%% Setup citation system
\usepackage[backend=biber, date=short, maxcitenames=2, bibencoding=utf8, style=apa]{biblatex}
\DeclareLanguageMapping{american}{american-apa}% needed for style=apa
%\renewcommand*{\bibfont}{\referencefont}
\setlength{\bibitemsep}{\baselineskip}
\setlength{\bibhang}{0pt}

\addbibresource{references.bib}% file with reference library

\usepackage{tikz}
\usetikzlibrary{fit,matrix,backgrounds}

\graphicspath{{graphics/}{Graphics/}{./}}%You can store your images in these locations

\title{Designing Authentic Industry-Engaged Assessment for Professional Competence in Business Intelligence}

\footer{Proceedings of the $22^\text{nd}$ International CDIO Conference, hosted by\\University of Liverpool, Liverpool, United Kingdom, June 22--25, 2026}

\setuphyperref{}%enable clickable links
\begin{document}
\makecdiotitle{
  \author{Helga Ingimundardóttir}
  \affiliation{Industrial Engineering, University of Iceland}
  % https://apastyle.apa.org/style-grammar-guidelines/capitalization/title-case
}

\section{Abstract}
Engineering programs increasingly seek to align student learning with professional practice, yet meaningful industry collaboration often exposes misalignment between academic assessment norms and workplace expectations. This paper presents a CDIO Implementation case describing the redesign of an undergraduate business intelligence course to strengthen industrial relevance while supporting student professional development. The redesign integrated semester-long, industry-inspired project work structured around team-based learning and competence-oriented assessment, emphasizing authentic data, reproducible workflows, and professional collaboration tools. A single industrial partner contributed real-world datasets, problem framing, and feedback, enabling students to engage with constraints and quality standards characteristic of professional practice. Assessment and feedback were redesigned to balance team performance with individual contribution through visible artifacts (e.g. version control activity) and structured reflection, addressing persistent challenges related to workload, accountability, and fairness. Evidence from iterative course evaluations, student reflections, graduate feedback, and industry partner input indicates increased student engagement, improved coherence across learning activities, and clearer alignment between assessment criteria and professional expectations. At the same time, the collaboration revealed tensions between academic tolerance for developmental errors and industry demands for reliability and trustworthiness, highlighting the importance of explicitly preparing students to communicate differently with academic and professional audiences. The paper reflects on these tensions and outlines design principles for industry-engaged CDIO courses, including expectation alignment early in the course, sustained use of a shared industrial context, and assessment practices that foreground professional standards without sacrificing formative learning space. The case offers transferable insights for CDIO programs seeking to leverage industrial collaboration as a catalyst for student professional development through innovative assessment and feedback.

\section{Keywords}
Industry collaboration, professional competence,
assessment and feedback, team-based learning, CDIO Standards: 5, 7, 8, 11


% Your paper directly addresses:
% ✅ Professional skills development
% ✅ Authentic, work-integrated assessment
% ✅ Feedback as a learning process
% ✅ Alignment between learning outcomes, activities, and assessment
% ✅ Bridging academia and industry

%From a CDIO perspective, your paper contributes particularly to:
%CDIO Standard 8 – Active Learning
%CDIO Standard 11 – Learning Assessment
%CDIO Standard 7 – Integrated Learning Experiences
%CDIO Standard 5 – Design-Implement Experiences

%%%%%%% DO NOT REMOVE %%%%%%%%%%%%%%%
\newpage{}% Title, authors, abstract, and keywords go on the first page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{body}

\section{ACKNOWLEDGEMENTS}
The author received no financial support for this work. This study was initially developed as a reflective case as part of an application to the Teaching Academy of the public universities in Iceland. The author would like to thank the Teaching Academy review committee for their constructive discussions and encouragement, which motivated the further development and dissemination of this work. The author also gratefully acknowledges the collaboration with Tixly Ticketing for contributing real-world data and professional perspectives that supported the authenticity of the course projects, as well as the students whose engagement informed the iterative refinement of the course design.

%\section{REFERENCES}
\setlength\bibitemsep{2pt}% Word template is single spaced, but it looks like there is extra space between items in the official implementation --foley

\printbibliography{}

\section{Biographical Information}
\textbf{\textit{Helga Ingimundardóttir}} is an Assistant Professor of Industrial Engineering at the University of Iceland. Since 2023, her teaching and research have focused on optimization, mathematical modeling, data science, artificial intelligence, and industry-engaged learning in data-intensive contexts. She is a member of the Teaching Academy of the public universities in Iceland (2024).
Helga brings extensive industry experience across research and applied AI: as a Research Scientist at \emph{deCODE Genetics}, she designed long-read sequencing analysis pipelines and co-authored papers in \emph{Nature Genetics} and \emph{Genome Biology}; as a Data Scientist at \emph{CCP Games}, she helped develop real-time recommendation services for \emph{EVE Online}; as Head of AI Research at \emph{Travelshift}, she led large-scale trip-planning optimization; and earlier, as a Computational Engineer in R\&D at \emph{Valka} (now JBT Marel), she developed X-ray-based optimization methods for automated fish processing.
She holds a PhD and MSc in Computational Engineering and a BSc in Mathematics from the University of Iceland, as well as a graduate diploma in Teaching Studies for Higher Education.

\subsection{Corresponding author}
\CorrespondingAuthorBlock{
Helga Ingimundardóttir\\
University of Iceland\\
School of Engineering and Natural Sciences\\
Faculty of Industrial Engineering, Mechanical Engineering and Computer Science\\
VR-II, Hjarðarhagi 6\\
107 Reykjavík, ICELAND\\}
{helgaingim@hi.is}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: xetex
%%% End:
